---
title: "'Won't you be my (distance-based nearest) neighbor?' Using Simultaneous Autoregressive Models to Understand Housing Prices in the 55105"
author: "Alyssa Bulatek, Sam Liberman, and Emma Heth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document: default
  html_document: default
bibliography: Library.bib
abstract: "Here is an astract with a quick summary of why this paper is important and what are the main conclusions."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warnings=FALSE)
```

```{r, echo=FALSE}
library(sf)
library(stringr)
library(lubridate)
library(dplyr)
library(ggplot2)
library(maptools)
library(ape)
library(tidyr)
library(spdep)
# Load in data from .RData files
load('RamseyCountyDataAreal.RData')
load('RamseyCountyDataPoint.RData')
# Add indicator variables correctly for houses on the river, on summit, or that are governor's mansion
RamseyCountyDataAreal <- RamseyCountyDataAreal %>%
    mutate(RIVER = ifelse(str_detect(ST_NAME, 'Mississippi'), 'yes', 'no')) %>%
    mutate(SUMMIT = ifelse(str_detect(ST_NAME, 'Summit') & ANUMBER != 1006, 'yes', 'no')) %>%
    mutate(GOVERNOR = ifelse(ANUMBER == 1006, 'yes', 'no'))
RamseyCountyDataPoint <- RamseyCountyDataPoint %>%
    mutate(RIVER = ifelse(str_detect(ST_NAME, 'Mississippi'), 'yes', 'no')) %>%
    mutate(SUMMIT = ifelse(str_detect(ST_NAME, 'Summit') & ANUMBER != 1006, 'yes', 'no')) %>%
    mutate(GOVERNOR = ifelse(ANUMBER == 1006, 'yes', 'no'))
Ramsey55105Areal <- RamseyCountyDataAreal %>%
    filter(ZIP == '55105')
Ramsey55105Point <- RamseyCountyDataPoint %>%
    filter(ZIP == '55105')
```

```{r, echo = FALSE}
RamseyCountyDataPoint_plot <- RamseyCountyDataPoint %>%
    filter(GARAGESQFT > 0 &
           ACRES_DEED < 500 &
           EMV_TOTAL < 2500000)
```

```{r, echo=FALSE}
ggplot(Ramsey55105Areal) + 
    geom_sf(aes(fill = log10(EMV_TOTAL)), lwd = 0) +
    coord_sf(datum=st_crs(Ramsey55105Areal, value = 3857)) + 
    labs(fill="Logarithm of\ntotal estimated\nmarket value (USD)") +
    theme_grey(base_size = 16) 
#midpoint of legend is $339,800, which is the median estimated total market value in 55105.
```

```{r, echo=FALSE}
ggplot(Ramsey55105Areal) + 
    geom_sf(aes(fill = FIN_SQ_FT^0.305), lwd = 0) +
    coord_sf(datum=st_crs(Ramsey55105Areal, value = 3857)) + 
    scale_fill_gradientn(colours=c("#1D263B", "#5C6784","#A0EADE", "#4CB963","#157F1F"), values=c(0,0.07,0.105,0.2,1)^0.305,
                         guide="colourbar",
                         limits=c(0,14887)^0.305, breaks=c(0,1566,14887)^0.305,labels=c('459',"1,566","14,887"), 
                         name='Finished Square Footage') 
#midpoint of legend is 1566, which is the median finished square footage in 55105.
```

We found a positive correlation between both the finished square footage and garage square footage (with houses without garages removed) and the total estimated market value of the property. We also noticed a relationship between the acreage deeded to a property and the estimated market value.

```{r}
ggplot(data = RamseyCountyDataPoint_plot, mapping = aes(x = ACRES_DEED, y = EMV_TOTAL)) + geom_point() + 
    ggtitle("Total market value versus deeded acres") + xlab("Deeded acres") + ylab("Total estimated market value in dollars")
    # May want to remove some outliers of deeded acres
```

```{r}
ggplot(data = RamseyCountyDataPoint_plot, mapping = aes(x = FIN_SQ_FT, y = EMV_TOTAL)) + geom_point() + 
    ggtitle("Total market value versus finished square footage, sorted by use classification") + facet_wrap(~USECLASS1) + 
    xlab("Finished square footage") + ylab("Total estimated market value (USD)") +
    theme_grey(base_size = 16) 
```

Although we did not include any time series analysis in our project, we also examined the relationship between the decade the house was built and its estimated market value. Though there are many outliers for each decade, there seems to be an oscillating trend with time, with a local maximum during World War II and a large increase in the third quartile of housing value in the 2010s.

```{r}
ggplot(data = RamseyCountyDataPoint_plot, aes(x = DECADE_BUILT, group=DECADE_BUILT, y = EMV_TOTAL)) + geom_boxplot() + 
    ggtitle("Total market value versus decade built") + xlab("Decade built") + ylab("Total estimated market value in dollars")
    # May want to remove some outliers of total market value
```

```{r, echo=FALSE}
ggplot(Ramsey55105Areal) + 
    geom_sf(aes(fill = DECADE_BUILT), lwd = 0) +
    coord_sf(datum=st_crs(Ramsey55105Areal, value = 3857)) + 
    scale_fill_gradientn(colours=c("#1D263B", "#5C6784","#A0EADE", "#4CB963","#157F1F"), values=c(0,0.35,0.5,0.65,1),
                         guide="colourbar",
                         limits=c(1850,2010), breaks=c(1850,1930,2010),labels=c('1850',"1930","2010"), 
                         name='Date Built') 
#midpoint of legend is $325,300, which is the median value per unit in 55105.
```

```{r, echo = FALSE, results = "hide"}
# Distance-based method
DistNB <- dnearneigh(Ramsey55105Point,d1=0, d2=300) # d2 is in feet
plot(st_geometry(Ramsey55105Areal), border="grey") # Run this and the line below TOGETHER!
plot(DistNB,st_centroid(st_geometry(Ramsey55105Areal), of_largest_polygon), add=TRUE, col="blue", pch='.', cex=.1)
nb <- DistNB #Queen, Rook, KNB, DistNB
W <- nb2mat(nb, style='B', zero.policy = TRUE) 
W[1:10,1:10]
```

```{r, echo = FALSE, results = "hide"}
Moran.I(Ramsey55105Areal$EMV_TOTAL, W, na.rm=TRUE) #Moran's I of adjacency matrix W
RamseyLM <- lm(EMV_TOTAL ~ FIN_SQ_FT + GARAGESQFT + DECADE_BUILT + HOMESTEAD + USECLASS1 + RIVER + SUMMIT + GOVERNOR, data = Ramsey55105Areal) #create LM to model residuals using morans I 
summary(RamseyLM)
# Plot residuals against prediction
ggplot(data = Ramsey55105Areal) + geom_point(aes(x = predict(RamseyLM), y = RamseyLM$resid), alpha = 0.3) +
    geom_smooth(aes(x = predict(RamseyLM), y = RamseyLM$resid), method = 'lm', se = FALSE) + 
    xlab("Predicted estimated market value (USD)") + ylab("Residual") +
    theme_grey(base_size = 16) 
Moran.I(RamseyLM$residuals, W, na.rm=TRUE)
```

```{r, echo = FALSE, results = "hide", cache=TRUE}
library(spatialreg)
listW <- nb2listw(nb)
Dat_sar = spautolm(formula = EMV_TOTAL ~ FIN_SQ_FT + GARAGESQFT + DECADE_BUILT + HOMESTEAD + USECLASS1 + RIVER + SUMMIT + GOVERNOR, data = Ramsey55105Areal, listw = listW, family = "SAR")
summary(Dat_sar)
Ramsey55105Areal$sar_resid <- NA #there are some missing values in Y and X
Ramsey55105Areal$sar_resid <- resid(Dat_sar)
```

```{r}
qqnorm(Ramsey55105Areal$sar_resid)
qqline(Ramsey55105Areal$sar_resid)
Ramsey55105Areal  %>%
  ggplot()+
  geom_sf(aes(fill = abs(sar_resid)), lwd = 0) + #not logged to show outliers
  labs(fill="Residuals \n(estimated \nmarket value \n(USD))") +
  theme_grey(base_size = 16) 
Ramsey55105Areal  %>%
  ggplot()+
  geom_sf(aes(fill = log(abs(sar_resid))), lwd = 0) +     
  labs(fill="Logarithm of \nresiduals \n(estimated \nmarket value \n(USD))") +
  theme_grey(base_size = 16) #Map the residuals (logged to see variation for small residuals)
# There's no areas where the colors are really bright or really dark (clustered). There seems to be uniform color or variation across space. This indicates we did a good job of capturing the spatial correlation with the SAR model, or at least not over or underestimating home values in whole areas (systematically). There are a few outliers that the model doesn't predict well (the Governor's Mansion), but for the most part it accounts for spatial correlation well.
```

```{r, echo = FALSE, results = "hide"}
Moran.I(Ramsey55105Areal$sar_resid, W, na.rm=TRUE) #Are the residuals still spatially correlated?
```

```{r results = "hide", echo = FALSE}
# This is just sample code if we want to make a table:
#echo = FALSE forces it to not show the code
#results = "hide" hides the output but shows the code
require(xtable)
#Add examples
#https://cran.r-project.org/web/packages/xtable/vignettes/xtableGallery.pdf
```

```{r results='hide', echo = FALSE}
#matrixtable <- matrix(1:10,nrow=5)
#print(xtable(Dat_sar, caption='Captions goes within xtable'), comment=FALSE)
````


## Conclusions

Housing values in the 55105 neighborhood of St. Paul can be explained using the quantitative and categorical variables we examined as well as a spatial correlation component. Except for some outliers which have high housing values due to characteristics that are not captured in our initial linear modeling, our SAR model seems to explain the housing value data well. These outliers can also be accounted for in the SAR model.

** @dubin1999spatial states that if there is a systematic pattern in the residuals when mapped over space, there is still a correlation in the residuals that was not captured by the modeling process. However, we notice no strong patterns in the logarithm of the residuals. This indicates we did a good job of capturing the spatial correlation with the SAR model, or at least not over or underestimating home values in whole areas systematically. There are a few outliers that the model doesn't predict well (for instance, the Governor's Mansion and the properties along the Mississippi River), but for the most part it accounts for spatial correlation well.

This project had a few limitations in the research we conducted. We could not examine the demographic data of the homeowners/renters of 55105, which limited the scope of our research to focus on the clustering of expensive homes in Macalester-Groveland rather than access (or lack thereof) to affordable housing for people of lower socioeconomic status. Furthermore, we removed all apartment complexes from our analysis because the estimated total value of those buildings skewed our analysis, and we did not analyze the proximity of properties to different amenities that could be desirable (such as shops, parks, the Macalester campus, and the Mississippi river). Our analysis was hindered by computational efficiency, as the 300 foot neighborhoods we defined were smaller than we would have liked to create given the extent to which houses affect the value of others in the same area. However, if we had defined our neighborhoods to be any larger, our adjacency matrix $W$ would be too large to store on our computers.


## Acknowledgements

Thanks to Professors Brianna Heggeseth and Laura Smith for their support on this project. We do not thank the doors to Smail Gallery for opening and making life cold.

## References
